### 2015-07-22 AuthorInformationFile/author_information.tex
Submitting to JORS.

\documentclass{article}
\usepackage{fullpage}
\usepackage{parskip}

\title{Measuring the Price of Anarchy in Critical Care Unit Interactions\\
Author Information}
\author{Vincent Knight*
    \and
        Izabela Komenda
    \and
        Jeff Griffiths}

\begin{document}

\maketitle

\textbf{Author affiliations:}

\begin{itemize}
    \item Vincent Knight: Cardiff University, School of Mathematics
    \item Izabela Komdena: ABUHB Health Board, ABCi
    \item Jeff Griffiths: Cardiff University, School of Mathematics
\end{itemize}

\textbf{Corresponding author:}

Dr Vincent Knight\\
Cardiff School of Mathematics\\
Senghennydd Road,\\
Cardiff\\
CF24 4AG\\
(+44) 29 2087 5548\\
www.vknight.org\\
+Vincent Knight\\
@drvinceknight\\
Skype: drvinceknight

\textbf{Statement of contribution:}

This paper contributes to the operational research literature in two main ways:

\begin{itemize}
    \item  Application of Game Theory to a healthcare system with a view to
        understanding choice of performance targets.
    \item Theoretical results as to the existence of Nash equilibria in pure
        strategies for the specific game considered.
\end{itemize}

The contribution is novel by placing an analytical stochastic model of critical
care unit interaction in a game theoretic context. The utilities of the game are
outputs of the stochastic model. Further to this the game is solved using a
simple search thanks to a theoretic result proving the existence of a pure Nash
equilibrium.

Various numerical results are offered which help describe the best design of
incentives so that each critical care unit acts in a coordinated way.

\end{document}